---
title: "Fantasy Football Analysis"
author: "John Jurica"
date: "2023-11-03"
output: html_document
---

# Welcome to the fantasy football analysis page!
### I am doing an analysis of fantasy football data from our fantasy football league from 2014 through 2022

```{r load packages}

library(tidyverse)
library(readr)


```


Some data was formatted in Excel. Personal Identifiers were removed.


```{r read in the data}

teams_info_2022_thru_2014_updated <- read_csv("teams_info_2022_thru_2014_updated_csv.csv", 
  col_types = cols(
    `Team ID` = col_double(),               
    `Owner First Name` = col_character(),   
    `Is League Creator` = col_character(),  
    `Playoff Seed` = col_double(),          
    `Final Rank` = col_double(),            
    `Wins` = col_double(),                  
    `Losses` = col_double(),                
    `Ties` = col_double(),                  
    `Points For` = col_double(),            
    `Points Against` = col_double(),        
    `Year` = col_double()                   
  )
)

View(teams_info_2022_thru_2014_updated)
```

Standardize First Names

```{r standardizes first names}
teams_info_2022_thru_2014_updated <- teams_info_2022_thru_2014_updated %>%
  mutate(first_name_lower = ifelse(`Owner First Name` == "JD" , toupper(`Owner First Name`),
                                   ifelse(`Team ID` == 9, "Jimmy",  str_to_title(`Owner First Name`))))

View(teams_info_2022_thru_2014_updated)

```

Chart with Average Points For and Points Against

```{r Avg Points For and Against}
teams_info_2022_thru_2014_updated %>%
  group_by(`first_name_lower`) %>%
  summarise(avg_points_against = mean(`Points Against`), avg_points_for = mean(`Points For`)) %>%
  arrange(desc(avg_points_for))
```

Add a column to main dataset to indicate winner for ease of graphing

```{r Add a column for the winner each season}

teams_info_2022_thru_2014_updated <- 
  teams_info_2022_thru_2014_updated %>%
  mutate(winner_y_n = ifelse(`Final Rank` == 1, "yes", "no"))

View(teams_info_2022_thru_2014_updated)

```

Graph of Points For and Against per season. Removed player who only played a single season

```{r plots for points for and against}

scores_over_time_final <- teams_info_2022_thru_2014_updated %>%
  filter(first_name_lower != "Claude") %>% #removed single season player
  ggplot() +
  geom_point(aes(y = `Points For`, x = Year, shape = winner_y_n)) +
  geom_line(aes(y = `Points For`, x = Year), color = "blue") +
  geom_point(aes(y = `Points Against`, x = Year, shape = winner_y_n), color = "red") +
  geom_line(aes(y = `Points Against`, x = Year), color = "red") +
  labs(title = "Scores per year", subtitle = "Blue = Points For; Red = Points Against; Triangle = Winner", y = "Points") +
  facet_wrap(~first_name_lower) +
  theme(legend.position = "none")

scores_over_time_final 
```
